Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-10 csf-2017/paper.tex
changed how we write registers for instructions

\usepackage{rotating}

\usepackage{afterpage}

@@

% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\hs}{\var{ms}}
\newcommand{\ms}{\hs}
\newcommand{\hv}{\var{hv}}
\newcommand{\rv}{\var{rv}}
\newcommand{\lv}{\var{lv}}
\newcommand{\gl}{\var{g}}
\newcommand{\pc}{\mathit{pc}}
\newcommand{\pcreg}{\mathrm{pc}}
\newcommand{\addr}{\var{a}}
\newcommand{\offset}{\var{offset}}
\newcommand{\word}{\var{w}}
\newcommand{\start}{\var{base}}
\newcommand{\addrend}{\var{end}}
\newcommand{\pwlv}{\var{pwl}}
\newcommand{\mem}{\var{mem}}
\newcommand{\reg}{\var{reg}}
\newcommand{\heapseg}{\var{ms}}
\newcommand{\heap}{\var{mem}}
\newcommand{\mode}{\var{mode}}
\newcommand{\perm}{\var{perm}}
\newcommand{\permp}{\var{permPair}}
\newcommand{\roll}{\var{roll}}
\newcommand{\instr}{\var{instr}}
\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}
\newcommand{\adv}{\var{adv}}
\newcommand{\msframe}{ms_\var{frame}}
\newcommand{\link}{\var{link}}
\newcommand{\stk}{\var{stk}}
\newcommand{\flag}{\var{flag}}
\newcommand{\nwl}{\var{nwl}}
\newcommand{\pwl}{\var{pwl}}
\newcommand{\sta}{\var{sta}}
\newcommand{\cnst}{\var{cnst}}
\newcommand{\olf}{\var{offsetLinkFlag}}
\newcommand{\prp}{\var{prp}}
\newcommand{\env}{\var{env}}
\newcommand{\cls}{\var{cls}}
\newcommand{\unused}{\var{unused}}
\newcommand{\act}{\var{act}}

@@

% Reference register/memory
\newcommand{\refreg}[1]{\lfloor #1 \rfloor}
\newcommand{\refheap}[1]{\langle #1 \rangle_m}

@@

\begin{figure}
  \begin{align*}
    \addr   &\in& \Addrs &\isdef \nats\\
    w &\in&\Words &\isdef \ints + \Caps \\
    \perm   &\in& \Perms &::= \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid \exec\mid \entry\mid \rwx\mid \rwlx\\
    \gl&\in&\Globals & ::= \glob \mid \local \\
     && \Caps  &::= \left\{
                       \begin{multlined}
                         ((\perm,\gl),\start,\addrend,\addr)\mid\\
                         \start\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}
                       \end{multlined} \right\}\\
    r       &\in& \RegName&\bnfdef \pcreg\mid r_0\mid r_1\mid\ldots\\
    \reg &\in& \Regs  &\isdef \RegName \rightarrow \Words\\
    m&\in& \Heaps &\isdef \Addrs \rightarrow \Words \\
    \Phi    &\in& \ExecConfs  &\isdef \Regs \times \Heaps \\
    &&\Confs &::= \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps \\
    \ms     &\in& \MemSegments &::= \Addrs \parfun \Words
  \end{align*}
  \begin{equation*}
  \begin{array}{rcl}
    \lv    &::=& \refreg{r} \\
    \hv    &::=& \refheap{r}\\
    \rv    &::=& n \mid \lv \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid  \\
           &   & \plus{\lv}{\rv}{\rv} \mid 
                 \minus{\lv}{\rv}{\rv} \mid 
                 \lea{\lv}{\rv} \mid 
                 \restricttwo{\lv}{\rv} \mid 
                 \subseg{\lv}{\rv}{\rv} \mid  \\
           &   & \isptr{\lv}{\rv} \mid 
                 \getp{\lv}{\lv} \mid 
                 \getl{\lv}{\lv} \mid 
                 \getb{\lv}{\lv} \mid
                 \gete{\lv}{\lv} \mid\\
           &   & \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\end{equation*}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

@@

\begin{figure*}
  \centering
  \begin{equation*}
    \Phi  \rightarrow
    \begin{cases}
      \sem{\decode(n)}(\Phi) & \text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }\\
      \failed                                 & \text{otherwise}
    \end{cases}
  \end{equation*}\vspace{-.3cm}%
  \begin{align*}
    \updatePcPerm{w} &=
                       \begin{cases}
                         ((\exec,\gl),\start,\addrend,\addr) & \text{if $w = ((\entry,\gl),\start,\addrend,\addr)$}\\
                         w & \text{otherwise}
                       \end{cases}\\
    \stdUpdatePc{\Phi} &=
                         \begin{cases}
                           \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
                           \begin{array}[t]{l}
                             \text{if $\memreg(\pcreg) = \stdcap$}\\
                             \quad\text{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}\\
                           \end{array}
                         \end{cases}
  \end{align*}

  \begin{tabular}{|c|p{3.8cm}|p{10.4cm}|}
    \hline
    $i$&$\sem{i}(\Phi)$&Conditions\\
    \hline
    $\fail$&$\failed$&\\
    \hline
    $\halt$&$(\halted,\memheap)$&\\
    \hline
    $\move{\refreg{r_1}}{\rv}$& $\begin{multlined}
      \stdUpdatePc{}(\updateReg{r_1}{\\
          \memreg(\rv)}) \end{multlined}$&\\
    \hline
    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\updateReg{r_1}{\memheap(\addr)}$&$\memreg(r_2) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$  and  $\start \leq \addr \leq \addrend$\\
    \hline
    $\restricttwo{\refreg{r}}{\rv}$&$\begin{multlined}
                                       \stdUpdatePc{}(\updateReg{r}{\\(\permp',\start,\addrend,\addr)})
                                     \end{multlined}$  &
                                                        $\memreg(r) = \stdcap[\permp]$ and $\permp' = \decodePermPair{\memreg(\rv)}$
                                                      and $\permp'\sqsubseteq \permp$\\
    \hline
    $\geta{\refreg{r_1}}{\refreg{r_2}}$ & $\stdUpdatePc{\updateReg{r_1}{\addr}}$ &
                                                $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$\\
    \hline
    $\jmp{\lv}$&$\begin{multlined}
                   \updateReg{\pcreg}{\\\updatePcPerm{\memreg(\lv)}}
                 \end{multlined}$&\\
    \hline
    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and  $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
                                                                                       and if  $\var{w} = ((\_,\local),\_,\_,\_)$, then  $\perm \in \{\rwlx,\readwritel \}$\\
    \hline
    \multicolumn{3}{|c|}{$\cdots$}\\
    \hline
    \_&$\failed$&otherwise\\
    \hline
  \end{tabular}
  \caption{An excerpt from the operational semantics}
  \label{fig:op-sem}
\end{figure*}

>>>>

\usepackage{rotating}
\usepackage{afterpage}

@@

% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\hs}{\var{ms}}
\newcommand{\ms}{\hs}
\newcommand{\hv}{\var{r}}
\newcommand{\rv}{\var{r}}
\newcommand{\lv}{\var{r}}
\newcommand{\gl}{\var{g}}
\newcommand{\pc}{\mathit{pc}}
\newcommand{\pcreg}{\mathrm{pc}}
\newcommand{\addr}{\var{a}}
\newcommand{\offset}{\var{offset}}
\newcommand{\word}{\var{w}}
\newcommand{\start}{\var{base}}
\newcommand{\addrend}{\var{end}}
\newcommand{\pwlv}{\var{pwl}}
\newcommand{\mem}{\var{mem}}
\newcommand{\reg}{\var{reg}}
\newcommand{\heapseg}{\var{ms}}
\newcommand{\heap}{\var{mem}}
\newcommand{\mode}{\var{mode}}
\newcommand{\perm}{\var{perm}}
\newcommand{\permp}{\var{permPair}}
\newcommand{\roll}{\var{roll}}
\newcommand{\instr}{\var{instr}}
\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}
\newcommand{\adv}{\var{adv}}
\newcommand{\msframe}{ms_\var{frame}}
\newcommand{\link}{\var{link}}
\newcommand{\stk}{\var{stk}}
\newcommand{\flag}{\var{flag}}
\newcommand{\nwl}{\var{nwl}}
\newcommand{\pwl}{\var{pwl}}
\newcommand{\sta}{\var{sta}}
\newcommand{\cnst}{\var{cnst}}
\newcommand{\olf}{\var{offsetLinkFlag}}
\newcommand{\prp}{\var{prp}}
\newcommand{\env}{\var{env}}
\newcommand{\cls}{\var{cls}}
\newcommand{\unused}{\var{unused}}
\newcommand{\act}{\var{act}}

@@

% Reference register/memory
\newcommand{\refreg}[1]{#1}
\newcommand{\refheap}[1]{#1}

@@

\begin{figure}
  \begin{align*}
    \addr   &\in& \Addrs &\isdef \nats\\
    w &\in&\Words &\isdef \ints + \Caps \\
    \perm   &\in& \Perms &::= \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid \exec\mid \entry\mid \rwx\mid \rwlx\\
    \gl&\in&\Globals & ::= \glob \mid \local \\
     && \Caps  &::= \left\{
                       \begin{multlined}
                         ((\perm,\gl),\start,\addrend,\addr)\mid\\
                         \start\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}
                       \end{multlined} \right\}\\
    r       &\in& \RegName&\bnfdef \pcreg\mid r_0\mid r_1\mid\ldots\\
    \reg &\in& \Regs  &\isdef \RegName \rightarrow \Words\\
    m&\in& \Heaps &\isdef \Addrs \rightarrow \Words \\
    \Phi    &\in& \ExecConfs  &\isdef \Regs \times \Heaps \\
    &&\Confs &::= \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps \\
    \ms     &\in& \MemSegments &::= \Addrs \parfun \Words
  \end{align*}
  \begin{equation*}
  \begin{array}{rcl}
    \rv    &::=& n \mid \lv \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid  \\
           &   & \plus{\lv}{\rv}{\rv} \mid 
                 \minus{\lv}{\rv}{\rv} \mid 
                 \lea{\lv}{\rv} \mid 
                 \restricttwo{\lv}{\rv} \mid \\
           &   & \subseg{\lv}{\rv}{\rv} \mid  
                 \isptr{\lv}{\rv} \mid 
                 \getp{\lv}{\lv} \mid 
                 \getl{\lv}{\lv} \mid\\ 
           &   & \getb{\lv}{\lv} \mid
                 \gete{\lv}{\lv} \mid
                 \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\end{equation*}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

@@

\begin{figure*}
  \centering
  \begin{equation*}
    \Phi  \rightarrow
    \begin{cases}
      \sem{\decode(n)}(\Phi) & \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }}\\
      \failed                                 & \text{\footnotesize{otherwise}}
    \end{cases}
  \end{equation*}\vspace{-.3cm}%
  \begin{align*}
    \updatePcPerm{w} &=
                       \begin{cases}
                         ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
                         w & \text{\footnotesize{otherwise}}
                       \end{cases}\\
    \stdUpdatePc{\Phi} &=
                         \begin{cases}
                           \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
                             \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$ and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}}\\
                             \failed & \text{\footnotesize{otherwise}}
                         \end{cases}
  \end{align*}

  \begin{tabular}{|c|p{3.8cm}|p{10.4cm}|}
    \hline
    $i$&$\sem{i}(\Phi)$&Conditions\\
    \hline 
    $\fail$&$\failed$&\\
    \hline
    $\halt$&$(\halted,\memheap)$&\\
    \hline
    $\move{\refreg{r_1}}{r_2}$& $\begin{multlined}
      \stdUpdatePc{}(\updateReg{r_1}{\\
          \memreg(r_2)}) \end{multlined}$&\\
    \hline
    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\updateReg{r_1}{\memheap(\addr)}$&$\memreg(r_2) = \stdcap$\footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$\footnotesize{ and }$\start \leq \addr \leq \addrend$\\
    \hline
    $\restricttwo{\refreg{r_1}}{r_2}$&$\begin{multlined}
                                       \stdUpdatePc{}(\updateReg{r_1}{\\(\permp',\start,\addrend,\addr)})
                                     \end{multlined}$  &
                                                        $\memreg(r_2) = \stdcap[\permp]$\footnotesize{ and }\newline$\permp' = \decodePermPair{\memreg(r_2)}$
                                                      \footnotesize{ and }$\permp'\sqsubseteq \permp$\\
    \hline
    $\geta{\refreg{r_1}}{\refreg{r_2}}$ & $\stdUpdatePc{\updateReg{r_1}{\addr}}$ &
                                                $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$\\
    \hline
    $\jmp{\lv}$&$\begin{multlined}
                   \updateReg{\pcreg}{\\\updatePcPerm{\memreg(\lv)}}
                 \end{multlined}$&\\
    \hline
    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$\footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$ \footnotesize{ and } $\start \leq \addr \leq \addrend$\footnotesize{ and }$\var{w} = \memreg(r_2)$
                                                                                       \footnotesize{and if }$\var{w} = ((\_,\local),\_,\_,\_)$\footnotesize{, then  }$\perm \in \{\rwlx,\readwritel \}$\\
    \hline
    \multicolumn{3}{|c|}{$\cdots$}\\
    \hline
    \_&$\failed$&\footnotesize{otherwise}\\
    \hline
  \end{tabular}
  \caption{An excerpt from the operational semantics}
  \label{fig:op-sem}
\end{figure*}
